---
title: "EDH descriptive stats"
author: "Petra Hermankova"
date: "23/09/2020"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---



## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE}
devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
#devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
library(tidyverse)
library(sdam)
library(jsonlite)
library(leaflet)
library(tidytext)
library(psych)
```

## Loading data
1. Load the dataset, if you have Sciencedata.dk credentials

```{r, echo=FALSE}
mycred_secret<- readLines("~/mysecret.txt")
```

```{r, loading data}
resp = request("EDH_text_cleaned_2021-01-21.json", path="/sharingin/648597@au.dk/SDAM_root/SDAM_data/EDH/public", method="GET", cred=mycred_secret)
```

```{r, echo=FALSE}
remove(mycred_secret)
```


2. Make a list and tibble from the request function
```{r}
list_json <- jsonlite::fromJSON(resp)
EDH_tibble <- as_tibble(list_json)
```

3. Display the first 6 records
```{r}
head(EDH_tibble)
```

# What is in the dataset? Descriptive stats

## Total number of inscriptions
```{r}
nrow(EDH_tibble)
```


```{r}
# In June 2020 there were 80270 inscriptions. How many inscriptions there are now in total? What is the increse?
total<- nrow(EDH_tibble)
total
total - 80270
```

## Attributes

```{r}
names(EDH_tibble)
```



## Total number of inscriptions by language
```{r}
language<- as.data.frame(table(unlist(EDH_tibble$language)))
language

language<- language %>% 
  mutate(Percentage = Freq/(total/100))
language$Percent_round<- round(language$Percentage, digits = 2)
language$Percent_round
```

## Location

```{r}
location<- EDH_tibble %>% 
  select(id, coordinates, findspot_ancient, findspot_ancient_clean, findspot_ancient_certainty, province_label, province_label_clean, province_label_certainty, country, country_clean, country_certainty, modern_region, modern_region_clean, modern_region_certainty, findspot_modern, findspot_modern_clean, findspot_modern_certainty, findspot, findspot_clean, findspot_certainty, present_location, edh_geography_uri, commentary)
head(location)
```

### How many inscriptions have spatial coordinates
```{r}

# how many inscriptions don't have spatial coordinates
sum(str_count(location$coordinates, "list()"))

# how many % of isncriptions don't have spatial coordinates
sum(str_count(location$coordinates, "list()"))/(total/100)

# how many isncriptions have spatial coordinates available
total-sum(str_count(location$coordinates, "list()"))
                    
# how many % of isncriptions have spatial coordinates available
(total-sum(str_count(location$coordinates, "list()")))/(total/100)

```



#### Ancient findspot and the precision of coordinates

```{r}
location <- location %>% 
  separate(col = coordinates, into = c("latitude", "longitude"), sep = ",") %>% 
  mutate(latitude = str_extract(latitude, "[-]*\\d{1,2}\\.\\d+")) %>% 
  mutate(longitude = str_extract(longitude, "[-]*\\d{1,2}\\.\\d+"))
```

```{r}
decimalplaces<-function(x) {
  x_str <- toString(x)
  x_str<- str_split(x_str, "\\.")
  nchar(x_str[[1]][2])
}

### latitude

lat_df<- as.data.frame(location$latitude) %>%  
  mutate(decimals = unlist(map(location$latitude, decimalplaces)))

lat_df_count<- lat_df %>% 
  count(decimals, sort = T) %>% 
  mutate(percent = n/(sum(n)/100))

lat_df_count

### longitude

long_df<- as.data.frame(location$longitude) %>%  
  mutate(decimals = unlist(map(location$longitude, decimalplaces)))

long_df_count<- long_df %>% 
  count(decimals, sort = T) %>% 
  mutate(percent = n/(sum(n)/100))

long_df_count
```

### Known ancient findspot / `ancient_findspot_clean` attribute
```{r}
psych::describe(location$findspot_ancient_clean)
```

```{r}
# select ancient findspots where the value == NULL and group by the certainty
location %>% filter(findspot_ancient_clean == "NULL") %>%  count(findspot_ancient_certainty)
```

#### Uncertain ancient findspot (?)
```{r}
# how many inscriptions with Uncertain ancient findspot do we have?
location %>% filter(findspot_ancient_certainty == "Uncertain") %>% nrow()

location %>% filter(findspot_ancient_certainty == "Uncertain") %>% nrow()/(total/100)
```
 
```{r}
 # how many inscriptions with Uncertain ancient findspot grouped by the name of the ancient findspot 
location %>% filter(findspot_ancient_certainty == "Uncertain") %>%  count(findspot_ancient_clean, sort = TRUE)
```
 

### Information about province
```{r}
nrow(dplyr::filter(location, location$province_label_clean != "unknown"))
nrow(dplyr::filter(location, location$province_label_clean != "unknown"))/(total/100)

nrow(dplyr::filter(location, location$province_label_clean == "unknown"))
nrow(dplyr::filter(location, location$province_label_clean == "unknown"))/(total/100)
```

#### Uncertain Roman province
```{r}
nrow(dplyr::filter(location, location$province_label_certainty == "Uncertain"))/(total/100)
province_uncert<- dplyr::filter(location, location$province_label_certainty == "Uncertain")
province_uncert

# what is the present day location of inscriptions with an uncertain province
table(unlist(province_uncert$present_location))

# how many of them are from Rome
str_subset(province_uncert$present_location, "Roma") 
```

#### Unknown province
```{r}
nrow(dplyr::filter(location, location$province_label_clean == "unknown"))/(total/100)
```


### Inscriptions with no coordinates available
```{r}
no_coords <- location %>% 
  dplyr::filter(coordinates == "list()")
nrow(no_coords)

nrow(no_coords)/(total/100)
```

#### number of inscriptions with no coordinates but with ancient findspot known
```{r}
no_coords_findspot_anc <- no_coords %>% 
  filter(findspot_ancient != "list()")

nrow(no_coords_findspot_anc)
nrow(no_coords_findspot_anc)/(total/100)
```

#### number of inscriptions with no coordinates and ancient findspot unknown
```{r}
no_coords_no_findspot_anc <- no_coords %>% 
  filter(findspot_ancient == "list()")
nrow(no_coords_no_findspot_anc)
nrow(no_coords_no_findspot_anc)/(total/100)
```

#### number of inscriptions with no coordinates, unknwon ancient findspot but know modern findspot
```{r}
no_coords_findspot_mod <- no_coords_no_findspot_anc %>% 
  filter(findspot_modern != "list()")
nrow(no_coords_findspot_mod)
nrow(no_coords_findspot_mod)/(total/100)
```

#### number of inscriptions with no coordinates, no ancient nor modern findspot known
```{r}
no_coords_no_findspot_a_m <- no_coords_no_findspot_anc %>% 
  filter(findspot_modern == "list()")
nrow(no_coords_no_findspot_a_m)
nrow(no_coords_no_findspot_a_m)/(total/100)
```

#### inscriptions with no coordinates, no ancient nor modern findspot, but with known present location
```{r}
no_coords_present_location <- no_coords_no_findspot_a_m %>% 
  filter(present_location != "list()" & present_location != "lost")
nrow(no_coords_present_location)
nrow(no_coords_present_location)/(total/100)
```

##### How many inscriptions with only present day location are from museums or private collections in total numbers and percentages
```{r}
nrow(as.data.frame(str_subset(no_coords_present_location$present_location, "Mus|Muz")))
nrow(as.data.frame(str_subset(no_coords_present_location$present_location, "Mus|Muz")))/(nrow(no_coords_present_location)/100)
nrow(as.data.frame(str_subset(no_coords_present_location$present_location, "Mus|Muz")))/(total/100)


nrow(as.data.frame(str_subset(no_coords_present_location$present_location, "Priv")))
nrow(as.data.frame(str_subset(no_coords_present_location$present_location, "Priv")))/(nrow(no_coords_present_location)/100)
nrow(as.data.frame(str_subset(no_coords_present_location$present_location, "Priv")))/(total/100)
```


#### number of inscriptions with no coordinates, no ancient nor modern findpot known, nor their present location
```{r}
no_coords_no_findspot_a_m_no_present <- no_coords_no_findspot_a_m %>% 
  filter(present_location == "list()" | present_location == "lost")
nrow(no_coords_no_findspot_a_m_no_present)
nrow(no_coords_no_findspot_a_m_no_present)/(total/100)
```

#### inscriptions with no coordinates, no ancient or modern findspot or present location, only with province information
```{r}
no_coords_province<- no_coords_no_findspot_a_m_no_present %>% 
  filter(province_label_clean != "list()" & province_label_clean != "unknown")
nrow(no_coords_province)
nrow(no_coords_province)/(total/100)


no_coords_province_unknown <- no_coords_no_findspot_a_m_no_present %>% 
  filter(province_label_clean == "unknown")
nrow(no_coords_province_unknown)
```
#### inscriptions with no coordinates, no ancient or modern findspot or present location, only with modern country information
```{r}
no_coords_country<- no_coords_no_findspot_a_m_no_present %>% 
  filter(country != "list()" & country != "unknown")
nrow(no_coords_country)
nrow(no_coords_country)/(total/100)
```


#### inscriptions with no spatial information, not even province or modern country
```{r}
no_spatial_info<- no_coords_no_findspot_a_m_no_present %>% 
  filter(province_label_clean == "unknown" & country == "unknown")
nrow(no_spatial_info)
nrow(no_spatial_info)/(total/100)
```
##### inscriptions with no spatial info but with commentary addressing their origin

```{r}
unique(no_spatial_info$commentary)
```

### Checking the precision of spatial coordinates with individual spatial info (ancient findspot, modern findspot, present location etc) 
```{r}
# ancient findspot
coords_anc_findspot<- as.data.frame(cbind(id = location$id, coordinates = location$coordinates,findspot_ancient = location$findspot_ancient))

coords_anc_findspot<- coords_anc_findspot %>% 
  dplyr::filter(coordinates != "list()" & findspot_ancient != "list()")

nrow(coords_anc_findspot)/(total/100)


lat_long_anc_findspot<- coords_anc_findspot %>% 
  separate(col = coordinates, into = c("latitude", "longitude"), sep = ",")

lat_long_anc_findspot$latitude <- as.numeric(lat_long_anc_findspot$latitude)
lat_long_anc_findspot$longitude <- as.numeric(lat_long_anc_findspot$longitude)
```

### Creating a map with ancient findspots to check the precision of coordinates
```{r}
library(leaflet)
```

```{r}
map_ancient_findspot <-leaflet() %>%
  addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
  #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  addMarkers(lng = lat_long_anc_findspot$longitude, 
             lat = lat_long_anc_findspot$latitude,
             popup = paste0("<b> InscriptionID: </b>", lat_long_anc_findspot$id, 
                            "<br><b> Ancient Findspot: </b>", lat_long_anc_findspot$findspot_ancient),
    clusterOptions = markerClusterOptions()
  )
map_ancient_findspot
```

```{r}
map_all <-leaflet() %>%
  #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
  setView( lng = 20.326743, lat = 46.897122, zoom = 4 ) %>% 
  addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  addCircles(lng = EDH_tibble$longitude, 
             lat = EDH_tibble$latitude, opacity = 0.1, radius = 0.5, fill = TRUE, color = "orange",
             popup = paste0("<b> InscriptionID: </b>", EDH_tibble$id, 
                            "<br><b> Ancient Findspot: </b>", EDH_tibble$findspot_ancient_clean)
  ) %>% 
addLegend(position = "topright",
  colors = c("Orange"),
  labels = c("Findspots"), opacity = 1,
  title = "Spatial distribution of Latin inscriptions (EDH, n = 81,476)"
)
map_all
```

## Date

### How many % inscription has a start date
```{r}
not_before<- as.data.frame(unlist(EDH_tibble$not_before))

nrow(na.omit(not_before))
nrow(na.omit(not_before))/(total/100) 
```

### How many % inscription has an end date
```{r}
not_after<- as.data.frame(unlist(EDH_tibble$not_after))
nrow(na.omit(not_after))
nrow(na.omit(not_after))/(total/100)
```

### Earliest date
```{r}
describe(as.numeric(not_before$`unlist(EDH_tibble$not_before)`))
```

### Latest date
```{r}
describe(as.numeric(not_after$`unlist(EDH_tibble$not_after)`))
```

### Number of inscriptions having both not_before and not_after date
```{r}
has_both_dates<- EDH_tibble %>% 
  select(not_before, not_after) 

has_both_dates<- has_both_dates %>% 
  mutate(not_before = as.numeric(unlist(not_before))) %>% 
  mutate(not_after = as.numeric(unlist(not_after)))

has_both_dates <- has_both_dates %>% 
  filter(!is.na(not_before) & !is.na(not_after)) 
  
nrow(has_both_dates)

nrow(has_both_dates)/(nrow(EDH_tibble)/100)
```
### The length of dating interval
```{r}
has_both_dates<- has_both_dates %>% 
  mutate(date_interval = (not_before - not_after) * -1)

summary(has_both_dates$date_interval)
```


```{r}
has_both_dates %>% 
  count(date_interval, sort= T) %>% 
  mutate(perc = n/(sum(n)/100))
```


### Additional dating information available through the 'commentary'
```{r}
date_info <- EDH_tibble %>% 
  select(id, not_before, not_after, commentary, origdate_text, year_of_find) 

date_commentary <- date_info %>% 
  filter(commentary != "list()" )
date_commentary

nrow(date_commentary)
nrow(date_commentary)/(total/100)

str_subset(date_commentary$commentary, "Datierung")
```


```{r}
plot(table(EDH_tibble$not_before), ylab= "Number of inscriptions", xlab="'Not before' YEAR")
```


# Text of inscription

## How many inscriptions contain a text of an inscription

```{r}
length(na.omit(EDH_tibble$clean_text_interpretive_word))
```

In percent:
```{r}
length(na.omit(EDH_tibble$clean_text_interpretive_word))/(nrow(EDH_tibble)/100)
```

## How many words there are:

Original text before cleaning:
```{r}
sum(lengths(gregexpr("\\w+", EDH_tibble$transcription)) + 1)

#different counting method
sum(na.omit(str_count(EDH_tibble$transcription, '\\w+')))
```

Text after cleaning:
```{r}
sum(lengths(gregexpr("\\w+", EDH_tibble$clean_text_interpretive_word)) + 1)

#different counting method
sum(na.omit(str_count(EDH_tibble$clean_text_interpretive_word, '\\w+')))
```