---
title: "Exploring epigraphic formulae"
author: "Petra Hermankova"
date: "4/8/2021"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

## Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

#devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
#devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
library(tidyverse)
#library(sdam)
library(jsonlite)
library(leaflet)
```

## Loading data

```{r}
getwd()

list_json <- jsonlite::fromJSON("../data/EDH_text_cleaned_2021-01-21.json")
EDH_tibble <- as_tibble(list_json)
#dir.create("../figures")
```

Display the first 6 records
```{r}
head(EDH_tibble)
```

# Epigraphic formulae


## Dis Manibus 
Typically occurs on funerary inscriptions, invocation of the underworld gods.

Canonic Versions:

D M
D M S
Dis Manibus
Diis Manibus
Dis Manibus Sacrum
Diis Manibus Sacrum


```{r}
dis_manibus <- c("\\b[D|d]i{1,2}s\\b, \\b[M|m]anibus\\b", "\\b[S|s]acrum\\b", "\\b[D|d]\\b \\b[M|m]\\b", "\\b[D|d]\\b \\b[M|m]\\b \\b[S|s]\\b")

EDH_tibble %>% 
  filter(clean_text_interpretive_word %in% dis_manibus) %>% 
  dplyr::select(clean_text_interpretive_word)

EDH_tibble %>% 
  filter(clean_text_interpretive_word == str_subset(clean_text_interpretive_word, "anibus")) %>% 
  dplyr::select(clean_text_interpretive_word)
```

## Hic Situs/Sita Est

Typically occurs on funerary inscriptions, introducing the name of the deceased, marking his or hers resting place.

Canonic Versions:

H S E
Hic Situs Est
His Sita Est

H S S
Hic Siti Sunt

## Requiescit in pace (RIP)

Typically occurs on funerary inscriptions, introducing the name of the deceased who is perting in peace.

REQ
REQUISCIT
REQUIES

Hic requiescit
Hic requiescit in pace
Requievit 

full list https://www.trismegistos.org/abb/detail.php?id=64741

