---
title: "Exploration of the 'people' attribute in EDH dataset"
author: "Petra Hermankova"
date: "04/08/2021"
output:
  html_document:
    theme: united
    toc: yes
    toc_float: true
    number_sections: true
    toc_depth: 2
    df_print: paged
---

## Initial setup

## Setup of the environment:

```{r setup, echo=TRUE, message=FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

#devtools::install_github("sdam-au/sdam") # loading SDAM custom package, if not working try devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
#devtools::install_github("mplex/cedhar", subdir="pkg/sdam")
library(tidyverse)
#library(sdam)
library(jsonlite)
library(leaflet)
#install.packages("tidyverse")
```

## Loading data from Sciencedata.dk

```{r}
list_json <- jsonlite::fromJSON("https://sciencedata.dk/public/b6b6afdb969d378b70929e86e58ad975/EDH_text_cleaned_2021-01-21.json")
EDH_tibble <- as_tibble(list_json)

# download as a local copy
#download.file("https://sciencedata.dk/public/b6b6afdb969d378b70929e86#e58ad975/EDH_text_cleaned_2021-01-21.json", "../data/inscriptions.json")

```


## Loading data locally
```{r}
list_json <- jsonlite::fromJSON("../data/EDH_text_cleaned_2021-01-21.json")
EDH_tibble <- as_tibble(list_json)
#dir.create("../figures")
```

Display the first 6 records
```{r}
head(EDH_tibble)
```


# Exploration of 'People' in EDH

0. What are names of all attributes within the 'people' attribute
1. How many people are in total in the EDH database?
2. How many people there are per inscription (average, min, max)
3. What is the gender ratio of people on inscriptions? (male, female, NA)
4. What are the names of unique values in the 'status' attribute? 
5. What is the ratio of different statuses, e.g. slave vs freedman
6. How many inscriptions have ‘Age’ category?
7. What is the average age of people (years, months, days)
8. What is the origo of people, what is the origo on funerary inscriptions


When I have the pointers how to get the data out of 'people' I will be looking at the following specific cases:

Specific case (funerary inscriptions; attribute 'type_of_inscription_clean' == 'epitaph')
1. How many people are on funerary inscriptions (total, average, min, max)
2. What is the ratio of genders on funerary inscriptions (male, female, NA)
3. What is the age of people on funerary inscriptions (total number of inscriptions with age, average, min, max)
4. What is the average age of people on funerary inscriptions by province

Specific case (gender composition)
1. Ratio of men/women on different types of inscriptions (attribute 'type_of_inscription_clean')




```{r}
EDH_tibble$people[1:2]
```

```{r}
EDH_unnested<- EDH_tibble[1:10] %>% 
  dplyr::select(people) %>% 
  unnest_auto(people)
```

## What are names of all attributes within the 'people' attribute
```{r}
setdiff(names(EDH_unnested), names(EDH_tibble))
```

## How many people are in total in the EDH database?

One way through gender
```{r}
EDH_unnested %>% 
  count(gender, sort = TRUE) -> gender

sum(gender$n)
```

Second way through nrow
```{r}
nrow(EDH_unnested)
```


## How many people there are per inscription (average, min, max)
```{r}
summary(as.numeric(EDH_unnested$person_id))
```
# Gender

## What is the gender ratio of people on inscriptions? (male, female, NA)
```{r}
EDH_unnested %>% 
  count(gender, sort = TRUE)
```

# Status

## What are the names of unique values in the 'status' attribute?
```{r}
EDH_unnested$status %>% 
  unique()
```


## What is the ratio of different statuses, e.g. slave vs freedman

```{r}
str_split_fixed(EDH_unnested$status, ";", n=3) %>% 
  as.data.frame() -> status

status %>% 
  cbind(combined = c(status$V1,status$V2,status$V3)) %>% 
  filter(combined != "") %>% 
  mutate(combined_clean = str_replace_all(string = combined, pattern = "\\?", replacement = "")) %>% 
  mutate(combined_clean = str_replace_all(string = combined_clean, pattern = "^ ", replacement = "")) %>% 
  count(combined_clean, sort=TRUE) -> status_counts

status_counts 
```

```{r}
status_counts %>% 
  mutate(combined_clean = reorder(combined_clean, n)) %>% 
  ggplot(aes(y=combined_clean, x=n, fill=combined_clean)) +
  geom_col(width=0.8, stat="identity") +
  coord_cartesian(xlim=c(0,10000)) +
  labs(x = "Number of instances", y = "Status category", title = "Overview of status references in the EDH dataset", subtitle = ggtitle(paste("n =", nrow(EDH_tibble), "inscriptions"))) +
  geom_label(aes(label= n)) +
  theme_linedraw(base_size = 12) +
  theme_minimal() 

ggsave("../figures/Status_overview.png", width = 12, height = 8) 
```
# Age

## How many inscriptions have ‘Age’ category?

```{r}
EDH_unnested %>% 
  dplyr::select('age: days', 'age: months', 'age: hours', 'age: years') %>%
  filter(!is.na(EDH_unnested$`age: years`) | !is.na(EDH_unnested$`age: months`) | !is.na(EDH_unnested$`age: days`) |!is.na(EDH_unnested$`age: hours`))
```


## What are the unique values for years
```{r}
unique(EDH_unnested$`age: years`)
```

## How many people have their age stated in years
```{r}
sum(!is.na(EDH_unnested$`age: years`))
EDH_unnested %>% 
  dplyr::select('age: days', 'age: months', 'age: hours', 'age: years') %>%
  filter(!is.na(EDH_unnested$`age: years`))
```

## How many people have their age stated in months
```{r}
sum(!is.na(EDH_unnested$`age: months`))
EDH_unnested %>% 
  dplyr::select('age: days', 'age: months', 'age: hours', 'age: years') %>%
  filter(!is.na(EDH_unnested$`age: months`))
```

## How many people have their age stated in days
```{r}
sum(!is.na(EDH_unnested$`age: days`))
EDH_unnested %>% 
  dplyr::select('age: days', 'age: months', 'age: hours', 'age: years') %>%
  filter(!is.na(EDH_unnested$`age: days`))
```

## How many people have their age stated in hours
```{r}
sum(!is.na(EDH_unnested$`age: hours`))
EDH_unnested %>% 
  dplyr::select('age: days', 'age: months', 'age: hours', 'age: years') %>%
  filter(!is.na(EDH_unnested$`age: hours`))
```


## What is the average age of people (years, months, days)

Not ideal method as it skips a lot of textual descriptions
```{r}
summary(as.numeric(EDH_unnested$`age: years`))
summary(as.numeric(EDH_unnested$`age: months`))
summary(as.numeric(EDH_unnested$`age: days`))
summary(as.numeric(EDH_unnested$`age: hours`))
```

Better method using regular expressions to detect years and converting them as numeric
```{r}
EDH_unnested %>% 
  dplyr::select('age: days', 'age: months', 'age: hours', 'age: years') %>% 
  mutate(age_years = as.numeric(str_extract(EDH_unnested$'age: years', pattern = "[:digit:]+"))) %>% 
  mutate(age_months = as.numeric(str_extract(EDH_unnested$'age: months', pattern = "[:digit:]+"))) %>%
  mutate(age_days = as.numeric(str_extract(EDH_unnested$'age: days', pattern = "[:digit:]+"))) %>%
  mutate(age_hours = as.numeric(str_extract(EDH_unnested$'age: hours', pattern = "[:digit:]+"))) -> ages
```

```{r}
summary(ages$age_years)
summary(ages$age_months)
summary(ages$age_days)
summary(ages$age_hours)
```

## Combining all ages (years, months, days, hours) into one column

```{r}

ages <- ages %>% 
  mutate(months_to_years = age_months / 12) %>% 
  mutate(days_to_years = age_days / 365) %>% 
  mutate(hours_to_years = age_hours / (24*365))


ages <- ages %>%
  replace_na(list(months_to_years = 0, days_to_years = 0, hours_to_years = 0)) %>% 
  mutate(total_age = age_years + months_to_years + days_to_years + hours_to_years) %>% 
  dplyr::select(-ends_with("to_years"))

```


## Combine the total age with the rest of the dataset

```{r}

EDH_age<- EDH_unnested %>% 
  mutate(age_years = as.numeric(str_extract(EDH_unnested$'age: years', pattern = "[:digit:]+"))) %>% 
  mutate(age_months = as.numeric(str_extract(EDH_unnested$'age: months', pattern = "[:digit:]+"))) %>%
  mutate(age_days = as.numeric(str_extract(EDH_unnested$'age: days', pattern = "[:digit:]+"))) %>%
  mutate(age_hours = as.numeric(str_extract(EDH_unnested$'age: hours', pattern = "[:digit:]+"))) %>% 
  mutate(months_to_years = age_months / 12) %>% 
  mutate(days_to_years = age_days / 365) %>% 
  mutate(hours_to_years = age_hours / (24*365)) %>% 
  replace_na(list(months_to_years = 0, days_to_years = 0, hours_to_years = 0)) %>% 
  mutate(total_age = age_years + months_to_years + days_to_years + hours_to_years) %>% 
  dplyr::select(-ends_with("to_years")) 

```

## Summary of age in years

```{r}
summary(EDH_age$total_age)
```

### How many percent of people state their age on inscriptions
```{r}
length(na.omit(EDH_age$total_age))/(nrow(EDH_age)/100)
```




# Occupation
## What are the names of unique values in the 'occupation' attribute?
```{r}
EDH_unnested %>% 
  dplyr::filter(occupation == "data available") %>% 
  select()
```
Unfortunately, ant other details about the occupation are not provided.

# Origin
```{r}
origo<- as.data.frame(EDH_unnested$origo) 
origo

origo %>% 
  filter(`EDH_unnested$origo` != "") %>% 
  #mutate(clean_origo = str_replace_all(string = combined, pattern = "\\?", replacement = "")) %>%
  count(`EDH_unnested$origo`, sort=TRUE)
```

## What kind of inscriptions has origo
```{r}
EDH_unnested %>% 
  filter(!is.na(origo)) %>% 
  count(type_of_inscription_clean, sort=T)

```
## What is the findspot vs origo
```{r}
EDH_unnested %>% 
  filter(!is.na(origo) & type_of_inscription_clean == "epitaph") %>% 
  select(origo, findspot_ancient_clean) %>% 
  count(findspot_ancient_clean, origo, sort=T)
```

## Cleaning origo

```{r}
EDH_unnested <- EDH_unnested %>%
  mutate(origo_clean = str_replace_all(origo, pattern="[\\*|\\+|\\?|\\!]", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="(\\w+) \\(= (\\w+)\\)", replacement="\\2")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="gente|genitus|gentis|natus|civis|Civis|civus|natione|nato|regione|domo|domu|cives|cive|civi|ex|verna|tate", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="^ ", replacement="")) %>%
  mutate(origo_clean = str_replace_all(origo_clean, pattern="  ", replacement=" ")) 

```

```{r}
EDH_unnested %>%
  select(origo_clean, id) -> EDH_origo_out
```


```{r}
write.csv(EDH_origo_out, "../data/EDH_origo_out.csv")
```

### Further cleaning in Open Refine, loading its results
```{r}
origo_OR<-read.csv("../data/EDH_origo_in_OR.csv")
```


```{r}
EDH_unnested <- cbind(EDH_unnested, origo_clean_OR = origo_OR$origo_clean)
```

### Saving the new dataset with clean origo to Sciencedata.dk

```{r}
#install.packages("Rtools")
#install.packages("devtools")
#install.packages("sdam")
#devtools::install_github("sdam-au/sdam")
library(sdam)
```

```{r}

mycred_secret<- readLines("../../../../Desktop/mysecret.txt")

EDH_unnested_json <- jsonlite::toJSON(EDH_unnested, auto_unbox = TRUE)
jsonlite::write_json(EDH_unnested_json, path="../data/EDH_people_clean_origo_2021-08-25.json")

sdam::request("../data/EDH_people_clean_origo_2021-08-25.json", path="/sharingout/648597@au.dk/SDAM_root/SDAM_data/EDH/public", force = TRUE,
        method="PUT", cred=c(mycred_secret[1], mycred_secret[2]))

remove(mycred_secret)
```


## Mapping inscriptions with origo

```{r}
EDH_unnested<- EDH_unnested %>% 
  separate(col = coordinates, into = c("longitude", "latitude"), sep = ",")

EDH_unnested$latitude <- as.numeric(str_replace(EDH_unnested$latitude, pattern = "\\)", replacement=""))
EDH_unnested$longitude <- as.numeric(str_replace(EDH_unnested$longitude, pattern = "c\\(", replacement=""))
```

```{r}
EDH_unnested %>% 
  filter(!is.na(origo)) -> has_origo
```

There are 2648 records that mention origo and have coordinates

```{r}
library(raster)
library(sf)

origo_mapped<- leaflet(width="100%") %>%
 #addProviderTiles("Stamen.Watercolor")%>% # Add CartoDB map tiles
 addProviderTiles("Stamen.TerrainBackground")%>% # Add CartoDB map tiles
 #addProviderTiles("Esri.WorldTopoMap", group = "Topo") %>%
 #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  #setView( lng = 35.9239625, lat = 31.9515694, zoom = 5 ) %>%
  #setMaxBounds(lat1=43.633977, lng1 =-11.227926 , lat2=35.133882 , lng2=50.882336) %>%
  #addPolylines(data = roads, color = "purple", weight = 1, opacity = 0.7) %>% 
 addCircles(lng = has_origo$longitude, 
             lat = has_origo$latitude, opacity = 0.1, radius = 2, fill = TRUE, color = "red" , fillColor = "red",
             ) %>% 
  #addCircles(lng = xx$X, 
  #           lat = xx$Y,
  #           opacity = 0.5, radius = 15, fill = TRUE, color = "red" , fillColor = "black", popup = paste0("<b> City: </b>", xx$Name)) %>% 
addLegend(position = "bottomright",
  colors = c("Red"),
  labels = c("Inscriptions"), opacity = 1,
  title = "Inscriptions with origo statement" 
) %>% 
  addScaleBar(position="bottomleft")

origo_mapped

```


## Heat map of inscriptions with origo

```{r}
# https://www.supplychaindataanalytics.com/leaflet-heatmaps-in-r/
 
#install.packages("leaflet.extras")
library(leaflet.extras)
library(RColorBrewer)

heat_origo <- has_origo %>% 
  leaflet(width="100%") %>%
  addTiles() %>% 
  #addProviderTiles("Esri.WorldImagery", group = "ESRI Aerial") %>%
  #addProviderTiles("Esri.WorldShadedRelief", group = "ESRI Aerial") %>% 
  #addProviderTiles("Stamen.TerrainBackground") %>%
  addProviderTiles("Stamen.TonerBackground") %>%
  #addProviderTiles("CartoDB.VoyagerNoLabels") %>%
  setView( lng = 12.9239625, lat = 41.9515694, zoom = 4 ) %>%
  #setMaxBounds(lat1=40.633977, lng1 =-4.227926 , lat2=35.133882 , lng2=40.882336) %>%
 
 addHeatmap(lng = ~as.numeric(na.omit(has_origo$longitude)), lat = ~as.numeric(na.omit(has_origo$latitude)), 
            intensity = 0.5, layerId = NULL, group = NULL, minOpacity = 0.5, #max = 1,
  radius = 2, blur = 2, gradient = "YlOrRd", cellSize = 1,
  )

heat_origo
```








# Special focus - Funerary inscriptions

## Age

```{r}
EDH_age %>% 
  filter(type_of_inscription_clean == "epitaph") -> epitaph
```

How many percent of people on funerary inscriptions state their age on inscriptions

```{r}
length(na.omit(epitaph$total_age))/(nrow(epitaph)/100)
```
### What is the age of people on funerary inscriptions by province
```{r, fig.height=8}
epitaph %>% 
  dplyr::select(total_age, province_label_clean) %>% 
  count(total_age, province_label_clean, sort=TRUE) %>% 
  ggplot(aes(x=total_age, y= fct_rev(province_label_clean))) + geom_point(alpha=0.5, color="darkblue") +
  theme_minimal()

ggsave("../figures/Age_years_provinces.png", width = 8, height = 8) 
```

I would like to see the inscription claiming 200 years of age. A typo? An ambitious guess?

### What is the average age of people on funerary inscriptions by province

```{r}
#install.packages("psych")
library(psych)

psych::describeBy(epitaph$total_age, group = epitaph$province_label_clean, mat = TRUE, digits = 2) -> age_provinces
head(age_provinces)
```

### Average age by provinces (All inscriptions)
```{r, fig.height=12}
age_provinces %>% 
  filter(n>0) %>% 
  mutate(group1 = reorder(group1, mean)) %>%  
  ggplot(aes(y=group1, x=mean)) +
  geom_col(color="white", fill="blue", width=0.8, stat="identity") +
  #coord_cartesian(xlim=c(0,80)) +
  theme_minimal() +
  theme(text = element_text(size=16)) +
  labs(y="Roman province", x="Years", title= "Average age of people on inscriptions in the EDH database by province", subtitle = "n= 7993 people")+ 
  #geom_label(aes(label= mean)) +
  geom_label(aes(label = mean), colour = "black", fontface = "bold", hjust = -0.1) 
  
ggsave("../figures/Age_average_years_provinces.png", width = 12, height = 12) 
```

How do you know it is 7993 records?

### ### Average age by provinces, with more than 100 inscriptions containing age per province
```{r, fig.height=12}
age_provinces %>% 
  filter(n>100) %>% 
  mutate(group1 = reorder(group1, mean)) %>%  
  ggplot(aes(y=group1, x=mean)) +
  geom_col(color="white", fill="blue", width=0.8, stat="identity") +
  #coord_cartesian(xlim=c(0,80)) +
  theme_minimal() +
  theme(text = element_text(size=16)) +
  labs(y="Roman province ", x="Years", title= "Average age of people on inscriptions in the EDH database by province with more than 100 inscriptions stating age", subtitle = "n= 7993 people")+ 
  #geom_label(aes(label= mean)) +
  geom_label(aes(label = mean), colour = "black", fontface = "bold", hjust = -0.1)
  
ggsave("../figures/Age_average_years_provinces_100plus_inscr.png", width = 12, height = 12) 
```


### Inscriptions with age by provinces
```{r, fig.height=12}
age_provinces %>% 
  filter(n>0) %>% 
  mutate(group1 = reorder(group1, n)) %>%  
  ggplot(aes(y=group1, x=n)) +
  geom_col(color="white", fill="purple", width=0.8, stat="identity") +
  #coord_cartesian(xlim=c(0,80)) +
  theme_minimal() +
  theme(text = element_text(size=16)) +
  labs(y="Roman province", x="n", title= "Instances of age information in the EDH database by province", subtitle = "n= 7993 people")+ 
  #geom_text(aes(label= n),hjust = -0.4) +
  geom_label(aes(label = n), colour = "black", fontface = "bold", hjust = -0.4) 

ggsave("../figures/Age_info_provinces.png", width = 12, height = 12) 
```

### Inscriptions by age groups

```{r}
EDH_age<- EDH_age %>%
  mutate(agegroup = case_when(total_age < 3 ~  "0-2.99",
                              total_age < 15 ~ "3-14.99",
                              total_age < 30 ~ "15-29.99",
                              total_age < 40 ~ "30-39.99",
                              total_age < 60 ~ "40-59.99",
                              total_age > 60 ~ "over 60"))

EDH_age$agegroup <- factor(EDH_age$agegroup, levels = c("0-2.99", "3-14.99","15-29.99", "30-39.99", "40-59.99","over 60"))  
```

# test works
```{r}
summary(EDH_age$total_age)
summary(EDH_age$agegroup)
```


```{r}
EDH_age %>% 
  filter(agegroup != "NA") %>%
  #count(agegroup, sort = TRUE) %>%
  #mutate(agegroup_sorted = reorder(agegroup, n)) %>% 
  ggplot() +
  geom_bar(mapping = aes(x = agegroup, fill = agegroup))+
  labs(x = "Age group (years)", y = "Number of individuals", title = "Representation of age groups on funerary inscriptions (EDH dataset)", subtitle = ggtitle(paste("n =", nrow(filter(EDH_age, agegroup != "NA")), "inscriptions")))
  #+ geom_label(aes(label = agegroup), colour = "black", fontface = "bold", hjust = -0.4)
  

  
ggsave("../figures/Age_groups_epitaphs.png", width = 12, height = 8) 
```
### Children younger than 10

```{r}
EDH_age<- EDH_age %>% 
  mutate(age10 = ifelse(total_age < 10, "under10", "over10"))
```

#### Province
```{r, fig.height=14, fig.width=10}
EDH_age %>% 
  count(age10, province_label_clean) %>% 
  ggplot(aes(fill=age10, y=province_label_clean, x=n)) +
  geom_bar(position="fill", stat="identity") +
  theme_minimal() +
  #theme(text = element_text(size=16)) +
  labs(y="Roman province", x="n", title= "Ratio of age children under 10 years on inscriptions per province") 
  #geom_text(aes(label= n),hjust = -0.4) 
  #geom_label(aes(label = n), colour = "black", fontface = "bold", hjust = -0.4) 

```

```{r}
EDH_age10<- EDH_age %>% 
  count(age10, province_label_clean, sort=F) %>% 
  spread(key=age10, value=n) 

EDH_under10<- EDH_age10 %>% 
  mutate(total = rowSums(EDH_age10[2:4], na.rm=TRUE)) %>% 
  mutate(under10_ratio = under10 / (total/100)) %>% 
  mutate(over10_ratio = over10 / (total/100)) %>% 
  mutate(age_stated = (over10+under10) / (total/100))


EDH_under10 
```
```{r, fig.height=14, fig.width=10}
EDH_under10 %>%
  dplyr::select(province_label_clean, over10_ratio) %>% 
  #sort(over10_ratio, decreasing = TRUE, na.last = TRUE) %>% 
  ggplot(aes(y=province_label_clean, x=over10_ratio)) +
  geom_bar(stat="identity", position = "dodge") +
  theme_minimal() +
  #theme(text = element_text(size=16)) +
  labs(y="Roman province", x="%", title= "Ratio of children under 10 years on inscriptions per province") 
  #geom_text(aes(label= n),hjust = -0.4) 
  #geom_label(aes(label = n), colour = "black", fontface = "bold", hjust = -0.4)
```

### Rounding of age on inscriptions
```{r}
EDH_age %>% 
  dplyr::select(total_age) %>% 
  filter(total_age == str_subset(total_age, "5$|0$")) -> rounded_age


all_ages <- as.data.frame(na.omit(EDH_age$total_age))
nrow(rounded_age)/(nrow(all_ages)/100)
```

```{r}
rounded_age %>% 
  count(total_age)
```
### Non-rounded ages

```{r, fig.width=10}
all_ages %>% 
  count(`na.omit(EDH_age$total_age)`) %>% 
  mutate(total_age =`na.omit(EDH_age$total_age)`) %>% 
  ggplot(aes(x=total_age, y =n, colour=-n)) +
  geom_point(alpha=0.5) +
  geom_smooth() +
  geom_vline(xintercept = 10, linetype="dotted", 
               color = "green", size=0.5) +
  geom_vline(xintercept = 30, linetype="dotted", 
                color = "blue", size=0.5) +
  geom_vline(xintercept = 50, linetype="dotted", 
                color = "red", size=0.5) +
  geom_vline(xintercept = 70, linetype="dotted", 
                color = "brown", size=0.5)
  
```
### Population pyramid in funerary inscriptions


Adapted from https://klein.uk/teaching/viz/datavis-pyramids/

```{r}
EDH_age %>% 
  dplyr::select(gender, total_age) %>% 
  na.omit(gender) %>% 
  na.omit(total_age)  %>% 
  mutate(gender = str_replace(gender, "M\\?", "male")) %>% 
  mutate(gender = str_replace(gender, "F\\?", "female")) -> population_age

## cut the age variable into age groups with 5-year intervals
population_age$age_cut <- cut(population_age$total_age, breaks = seq(0, 200, 5), right = FALSE) 
population_age$population <- 1 ## each sampled respondent represents 10 individuals

population_age <- population_age %>% 
  mutate(age = str_replace(age_cut, "\\[", "")) %>% 
  mutate(age = str_replace(age, "\\)", "")) %>% 
  mutate(age = str_replace(age, "\\,", " - ")) 

labels <- levels(population_age$age_cut) %>% 
  stringr::str_replace_all(.,
                                           c("\\[" = "",
                                             "\\)" = "",
                                             "\\," = " - "))
  
# Trying if setting age to factor will help with sorting
class(population_age$age)
population_age$age <- factor(population_age$age, levels=c(labels))

# As a back up: create a new column from desired age intervals that lists the group number to be used for sorting
population_age$order <- findInterval(population_age$total_age, c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,95,100, 500))


sort(unique(population_age$order))
head(population_age)
population_age %>% 
  filter(order == 4)

## aggregate the data by gender and age group
population_age <- aggregate(formula = population ~ gender + age + order, data = population_age, FUN = sum)

## sort data by first by gender, then by age groups using *order*
population_age <- with(population_age, population_age[order(order,gender),])

#labels
xabels<- c("0-4", "5-9", "10-14", "15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80-84", "85-89", "90-94", "95-99", "100+")
# labels_df <- as.data.frame(labels)
# breaks = seq(0, 160, 5)

# AS: the original is poorly annotated. None of the lines of code specified WHY it is run. 
# Errors or problematic parts not clearly marked

  

population_age <- population_age[,c("age","order", "gender","population")]

population_age$population <- ifelse(population_age$gender == "male", -1*population_age$population, population_age$population)

# It is incredible how the ggplot sort lexicographically despite the factor!!


population_age$orderedage <- reorder(population_age$age, sort(population_age$order)) #https://kohske.wordpress.com/2010/12/29/faq-how-to-order-the-factor-variables-in-ggplot2/


pyramid_age <- ggplot(population_age, aes(x = orderedage, y = population, fill = gender)) + 
  geom_bar(aes(order), data = subset(population_age, gender == "female"), stat = "identity") +
  geom_bar(aes(order), data = subset(population_age, gender == "male"), stat = "identity") + 
  #scale_y_continuous(labels = c("250", "0", "250")) + 
  xlab("Age groups")+
  scale_x_discrete(labels = xabels)+  # Still does not show!! But no Error. jeez.
  coord_flip()
pyramid_age
```

# Adela Different and Working (for change) Alternative
```{r}
EDH_age %>% 
  dplyr::select(gender, total_age) %>% 
  na.omit(gender) %>% 
  na.omit(total_age)  %>% 
  mutate(gender = str_replace(gender, "M\\?", "male")) %>% 
  mutate(gender = str_replace(gender, "F\\?", "female")) -> population_age


# The column plotted on x (and flipped) needs to be a factor, but the example above is not working, 
# so lets' cut and label data with data.table library
library(data.table)

agebreaks <- c(0,5,10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,95,100, 500)
agelabels <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34",
               "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
               "70-74","75-79","80-84","85-89","90-94", "95-99","100+")
## cut the age variable into age groups with 5-year intervals

setDT(population_age)[ , agegroups := cut(total_age, 
                                breaks = agebreaks, 
                                right = FALSE, 
                                labels = agelabels)]

head(population_age)

population_age$population <- 1 ## each sampled respondent represents 10 individuals

# aggregate the data by gender and age group
population_age <- aggregate(formula = population ~ gender + agegroups, data = population_age, FUN = sum)

## sort data by first by gender, then by age groups
population_age <- with(population_age, population_age[order(agegroups,gender),])


population_age <- population_age[,c("agegroups","gender","population")]

population_age$population <- ifelse(population_age$gender == "male", -1*population_age$population, population_age$population)

pyramid_age <- ggplot(population_age, aes(x = agegroups, y = population, fill = gender)) + 
  geom_bar(data = subset(population_age, gender == "female"), stat = "identity") +
  geom_bar(data = subset(population_age, gender == "male"), stat = "identity") + 
  #scale_y_continuous(labels = c("250", "0", "250")) + 
  coord_flip()
pyramid_age
```


### Average age by individual years 

Using not before and not after date separately.

```{r}
describeBy(epitaph$total_age, group = epitaph$not_before, mat = TRUE, digits = 2) -> age_not_before

describeBy(epitaph$total_age, group = epitaph$not_after, mat = TRUE, digits = 2) -> age_not_after
```

### When (which years) do people state their age on funerary inscriptions

```{r}
age_not_bf_plot<- age_not_before %>% 
  ggplot(aes(x=group1, y=n)) +
  geom_point() +
  geom_vline(xintercept = 1, linetype="dotted", 
                color = "green", size=0.5) +
  geom_vline(xintercept = 100, linetype="dotted", 
                color = "blue", size=0.5) +
  geom_vline(xintercept = 200, linetype="dotted", 
                color = "red", size=0.5)
```


```{r}
age_not_aft_plot<- age_not_after %>% 
  ggplot(aes(x=group1, y=n)) +
  geom_point() +
  geom_vline(xintercept = 1, linetype="dotted", 
                color = "green", size=0.5) +
  geom_vline(xintercept = 100, linetype="dotted", 
                color = "blue", size=0.5) +
  geom_vline(xintercept = 200, linetype="dotted", 
                color = "red", size=0.5)
```
Commentary: People state their age in the second century the most!

### What is the average age of people on inscriptions in time
```{r}
age_not_before %>% 
  filter(!is.na(mean)) %>% 
  ggplot(aes(x=as.numeric(group1), y=as.numeric(mean))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 0, linetype="dotted", 
               color = "green", size=0.5) +
  geom_vline(xintercept = 100, linetype="dotted", 
                color = "blue", size=0.5) +
  geom_vline(xintercept = 200, linetype="dotted", 
                color = "red", size=0.5) +
  geom_vline(xintercept = 300, linetype="dotted", 
                color = "brown", size=0.5)



```


```{r}
age_not_after %>% 
  filter(!is.na(mean)) %>% 
  ggplot(aes(x=as.numeric(group1), y=as.numeric(mean))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 0, linetype="dotted", 
               color = "green", size=0.5) +
  geom_vline(xintercept = 100, linetype="dotted", 
                color = "blue", size=0.5) +
  geom_vline(xintercept = 200, linetype="dotted", 
                color = "red", size=0.5) +
  geom_vline(xintercept = 300, linetype="dotted", 
                color = "brown", size=0.5)

```


## Gender

```{r}
epitaph %>% 
  count(gender, sort=TRUE) %>% 
  mutate(ratio_total = n/(nrow(epitaph)/100)) %>% 
  mutate(ratio_total = reorder(ratio_total, n)) %>%
  ggplot(aes(y=gender, x=n)) +
  geom_col(color="white", fill="orange", width=0.8, stat="identity", fill=gender) +
  coord_cartesian(xlim=c(0,30000)) +
  theme_minimal() +
  theme(text = element_text(size=14)) +
  labs(y="Gender category", x="Number of instances", title= "Gender ratio on epitaphs in the EDH database", subtitle = "n = 47,803 inscriptions" ) +
  geom_label(aes(label= n)) 
# +  geom_text(aes(label = n), colour = "red", fontface = "bold", hjust = -0.1) 
  
ggsave("../figures/Gender_total_epitaphs.png", width = 12, height = 12)

```

```{r}


epitaph %>% 
  count(gender, sort=TRUE) %>% 
  mutate(ratio_total = n/(nrow(epitaph)/100)) %>% 
  mutate(ratio_total = round(ratio_total, digits = 2)) %>% 
  filter(ratio_total >0.5) %>% 
  ggplot(aes(y=gender, x=ratio_total, fill=gender)) +
  geom_bar(width=0.8, stat="identity") +
  coord_cartesian(xlim=c(0,60)) +
  theme_minimal() +
  theme(text = element_text(size=16)) +
  labs(y="Gender category", x="%", title= "Gender ratio on epitaphs in the EDH database", subtitle = ggtitle(paste("n =", nrow(epitaph), "inscriptions" ))) +
  geom_label(aes(label= ratio_total), hjust = -0.1) 
#+ geom_text(aes(label = ratio_total)) 

ggsave("../figures/Gender_epitaphs.png", width = 8, height = 8)
```

```{r}
epitaph %>% 
  dplyr::filter(gender == "male"| gender == "M?") -> epitaph_male

epitaph %>% 
  dplyr::filter(gender == "female"| gender == "F?") -> epitaph_female
```

```{r}
gender_true_ratio <- as.data.frame(cbind(male=nrow(epitaph_male), female=nrow(epitaph_female)))
gender_true_ratio %>% 
  gather() %>% 
  rename(gender = key) %>% 
  rename(n = value) %>% 
  as.data.frame() -> gender_ratio 

gender_ratio %>% 
  ggplot(aes(x=gender, y=n)) +
  geom_col(color="white", fill="orange", width=0.5, stat="identity") +
  theme_minimal() +
  theme(text = element_text(size=14)) +
  labs(x="Gender category", y="Number of people", title= "Gender ratio on epitaphs in the EDH database", subtitle = "n = 47,803 inscriptions" ) + 
  #geom_label(aes(label= n)) +
  geom_label(aes(label = n), colour = "black", fontface = "bold", vjust = 0.5) 

ggsave("../figures/Gender_total_ratio_epitaphs.png", width = 12, height = 12)
  
```

```{r}
gender_ratio %>% 
  mutate(ratio = n/(sum(n))*100) %>%
  mutate(ratio = round(ratio, digits =2)) %>% 
  ggplot(aes(x=gender, y=ratio), color=gender) +
  geom_col(width=0.5, stat="identity", fill = "brown") +
  theme_minimal() +
  theme(text = element_text(size=14)) +
  labs(x="Gender category", y="%", title= "Gender ratio on epitaphs in the EDH database", subtitle = "n = 47,803 inscriptions" ) + 
  #geom_label(aes(label= ratio)) + 
  geom_label(aes(label = ratio), colour = "black", fontface = "bold", vjust = 0.5) 

ggsave("../figures/Gender_ratio_epitaphs.png", width = 12, height = 12)
```











